---
title: "Sulcal morphology of posteromedial cortex substantially differs between humans and chimpanzees"
author: "EH Willbrand et al."
date: "4/17/2023"
output: pdf_document
---

# Libraries
## Requirements: packages below must be installed to load.
```{r setup, include=FALSE}
# Set working directory to "morph_csv" folder
getwd()
#setwd('~/PosteromedialSulci_Chimpanzees/morph_csv')
#knitr::opts_knit$set(root.dir = '~/PosteromedialSulci_Chimpanzees/morph_csv')

# Uncomment and run code below if packages are not installed
# project_packages <- c("tidyverse", "ggpattern", "rstatix", "effectsize", 
#                       "nlme", "RColorBrewer", "RVAideMemoire", "car")
# install.packages(project_packages)

# Load packages
library(tidyverse)
library(ggpattern)
library(rstatix)
library(effectsize)
library(nlme)
library(RColorBrewer)
library(RVAideMemoire)
library(car)

# Set themes and options
options(scipen=999, # 0
        show.signif.stars = FALSE) 

theme_set(theme_classic())

plot_theme <- theme(axis.title.x = element_text(size=0),
                    axis.title.y = element_text(size=16),
                    axis.text.x = element_text(size = 14),
                    axis.text.y = element_text(size = 14),
                    legend.title = element_text(size=16), 
                    legend.text = element_text(size=14),
                    strip.text.x = element_text(size = 14),
                    strip.placement = "outside",
                    axis.line = element_line(colour = "black", 
                                             linetype = "solid", 
                                             lineend = "round")
                    )

plot_theme2 <- theme(axis.title.x = element_blank(),
                    axis.title.y = element_text(size=16),
                    axis.text.x = element_text(size = 14),
                    axis.text.y = element_text(size = 14),
                    legend.title = element_text(size=16), 
                    legend.text = element_text(size=14),
                    strip.text.x = element_blank(),
                    strip.placement = "outside",
                    strip.background = element_blank(),
                    axis.line = element_line(colour = "black", 
                                             linetype = "solid", 
                                             lineend = "round")
                    )

plot_theme3 <- theme(axis.title.x = element_text(size=16),
                    axis.title.y = element_text(size=16),
                    axis.text.x = element_text(size = 14),
                    axis.text.y = element_text(size = 14),
                    legend.title = element_text(size=16), 
                    legend.text = element_text(size=14),
                    strip.text.x = element_text(size = 14),
                    strip.placement = "outside",
                    strip.background = element_blank(),
                    axis.line = element_line(colour = "black", 
                                             linetype = "solid", 
                                             lineend = "round")
                    )
```


# Main dataframes
```{r}
# PMC morphology
## Incidence
human_sulc_pmc <- read.csv("PMC_ROI_sulc_pct_human_72_subs.csv")
chimp_sulc_pmc <- read.csv("PMC_ROI_sulc_pct_chimp_60_subs.csv")
human_sulc_pmc$species <- "human"
chimp_sulc_pmc$species <- "chimpanzee"

sulc_pmc <- rbind(human_sulc_pmc, chimp_sulc_pmc)

### Morphology
human_morphology <- read.csv("human_pmc_morphology_final.csv")
chimp_morphology <- read.csv("chimp_pmc_morphology_final.csv")
human_morphology$species <- "human"
chimp_morphology$species <- "chimpanzee"

pmc_species_morphology <- rbind(human_morphology, chimp_morphology)


## Mcgs-specifc dataframes
### Incidence
mcgs_patterns <- read.csv("mcgs_patterns_final.csv")

### Morphology
human_mcgs_morphlogy <- read.csv("human_mcgs_morphology_final.csv")
chimp_mcgs_morphlogy <- read.csv("chimp_mcgs_morphlogy_final.csv")
human_mcgs_morphlogy$species <- "human"
chimp_mcgs_morphlogy$species <- "chimpanzee"

mcgs_species_morphology <- rbind(human_mcgs_morphlogy, chimp_mcgs_morphlogy)


## Data for normalizing 
lobe_morphology <- read.csv("lobe_morphology.csv")
insula_depth <- read.csv("insula_depth.csv")
```


# Analysis 1: Comparing percent of PMC in sulci in chimpanzees versus humans
1. We ran a linear mixed effects model (LME) with predictors of hemisphere and species, as well as their interaction, for percent overlap.
* Percent overlap was quantified as the DICE coefficient — overlap between the PMC as defined by the Destrieux parcellation and the FreeSurfer .sulc file (thresholded at > 0 for sulci).
* Hemisphere and species were considered fixed effects. 
* Hemisphere was nested within subjects. 
* We also controlled for differences in brain size in the model (quantified as the total cortical surface area of the given hemisphere). 

## Statistics
```{r}
# Merge dataframes
sulc_pmc.1 <- merge(sulc_pmc, lobe_morphology, by = c("sub", "hemi", "species"))


# Analysis 1

## Run analysis
model_pct_sulc_species <- lme(dice_coeff ~ hemi*species + cortex_sa, 
                              random = (~1|sub/hemi), 
                              data = sulc_pmc.1)

## View summary
model_pct_sulc_species.aov <- anova(model_pct_sulc_species)
model_pct_sulc_species.aov

## Effect size
effectsize::eta_squared(model_pct_sulc_species.aov)

## Digest main effect
sulc_pmc %>% group_by(species) %>% summarise(mean = mean(dice_coeff),
                                                   sd = sd(dice_coeff))
```

## Plot (Figure 1b)
```{r}
# Figure 1b

# Specific subject data
sulc_pmc.sub.c20.rh <- sulc_pmc %>% subset(sub %in% c("c20") & 
                                             hemi == "rh")
sulc_pmc.sub.151829.rh <- sulc_pmc %>% subset(sub %in% c("151829") & 
                                                hemi == "rh")

# Make plot 
sulc_pmc.plot <- sulc_pmc %>% 
  ggplot(aes(x = species, y = dice_coeff*100)) + 
  
  geom_violin(aes(fill = species), alpha = .5) +

  geom_boxplot(aes(fill = species), alpha = .5,
               outlier.size = 0, 
               outlier.alpha = 0, 
               width = .13) +
  
  geom_point(data = sulc_pmc.sub.151829.rh, 
             shape = 21, fill = "#d95f02", size = 2) +

  geom_point(data = sulc_pmc.sub.nc26.rh, 
             shape = 21, fill = "#e7298a", size = 2) +
  
  scale_y_continuous(name = "% sulci", 
                     limits = c(50, 90), n.breaks = 5) +
  
  facet_wrap(~ hemi, ncol = 2) + 
  
  scale_fill_manual(breaks = c("human", "chimpanzee"),
                    values = c("#252525", "#cccccc")
                    ) +
  
  guides(fill = "none") +

  plot_theme2

# View plot
sulc_pmc.plot

# Save plot
# ggplot2::ggsave(filename = "Figure_1b_sulc_pmc_plot.png",
#                  plot = sulc_pmc.plot,
#                  device = "png",
#                  width = 8,
#                  height = 5,
#                  units = "in",
#                  dpi = "retina")
```


# Analysis 2: Comparing incidence of PMC sulci in chimpanzees versus humans
1. We first tested for broad differences in sulcal incidence in PrC and PCC separately using LMEs with predictors of hemisphere and species, as well as their interaction, for sulcal count. 
* Hemisphere and species were considered fixed effects. 
* Hemisphere was nested within subjects. 

2. Next, for any sulcus that was not present in all hemispheres for either species, we tested the influence of species and hemisphere on the probability of a sulcus to be present with binomial logistic regression GLMs.
* For each statistical model, species (human, chimpanzee) and hemisphere (left, right), as well as their interaction, were included as factors for presence [0 (absent), 1 (present)] of a sulcus.

3. Finally, we conducted a follow-up GLM with PCC region [dorsal PCC (sspls-d, ifrms, icgs-p), ventral PCC (isms, sspls-v)], species (human, chimpanzee), and hemisphere (left, right), as well as their interaction, as factors for the presence of a sulcus.

## Statistics
```{r}
# Set up data for analyses
pmc_species_morphology.chisq <- pmc_species_morphology %>% mutate(pres_abs2 = case_when(pres_abs == "absent" ~ 0,
                                                                                   pres_abs == "present" ~ 1))

pmc_species_morphology.chisq$species <- as.factor(pmc_species_morphology.chisq$species)
pmc_species_morphology.chisq$hemi <- as.factor(pmc_species_morphology.chisq$hemi)

pmc_species_morphology.chisq.subject.prc <- pmc_species_morphology.chisq %>% 
  drop_na() %>%
  subset(label %in% c( "prculs_d", "prculs_v", "prcus_a", "prcus_i", "prcus_p")) %>% 
  group_by(sub, species, hemi) %>% 
  summarise(n = n())

pmc_species_morphology.chisq.subject.pcc <- pmc_species_morphology.chisq %>% 
  drop_na() %>%
  subset(label %in% c("icgs_p", "ifrms", "isms",  "sspls_d", "sspls_v")) %>% 
  group_by(sub, species, hemi) %>% 
  summarise(n = n())

column1 <- c("c14", "c15", "c18", "c23", "c26", 
             "c30", "c30", 
             "nc04", "nc05", "nc07", 
             "nc08", "nc12", "nc19", "nc32")
column2 <- c("chimpanzee", "chimpanzee", "chimpanzee", "chimpanzee", "chimpanzee", 
             "chimpanzee", "chimpanzee", "chimpanzee", "chimpanzee", "chimpanzee", 
             "chimpanzee", "chimpanzee", "chimpanzee", "chimpanzee")
column3 <- c("lh", "lh", "lh", "rh", "rh", 
             "lh", "rh",
             "rh", "lh", "rh", 
             "rh", "lh", "lh", "rh")
column4 <- c(0, 0, 0, 0, 0, 
             0, 0, 0, 0, 0, 
             0, 0, 0, 0)

chimps_no_pcc_sulci.df <- data.frame(column1, column2, column3, column4)
chimps_no_pcc_sulci.df <- chimps_no_pcc_sulci.df %>% 
  rename(sub = column1, species = column2, hemi = column3, n = column4)
pmc_species_morphology.chisq.subject.pcc <- rbind(pmc_species_morphology.chisq.subject.pcc, chimps_no_pcc_sulci.df)


# Analysis 2.1

## PrC sulci count

### Run analysis
model_num_sulc_prc <- lme(n ~ species * hemi, random = (~1|sub/hemi), 
                          data = pmc_species_morphology.chisq.subject.prc)

### View summary
model_num_sulc_prc.aov <- anova(model_num_sulc_prc)
model_num_sulc_prc.aov

### Effect size
effectsize::eta_squared(model_num_sulc_prc.aov)


## PCC sulci count

### Run analysis
model_num_sulc_pcc <- lme(n ~ species * hemi, random = (~1|sub/hemi), 
                          data = pmc_species_morphology.chisq.subject.pcc)

### View summary
model_num_sulc_pcc.aov <- anova(model_num_sulc_pcc)
model_num_sulc_pcc.aov

### Effect size
effectsize::eta_squared(model_num_sulc_pcc.aov)


# Analysis 2.2

## prculs-d
### Run analysis
pmc_species_morphology.chisq.prculs_d <- pmc_species_morphology.chisq %>% subset(label == "prculs_d")
pmc_species_morphology.chisq.prculs_d.glm <- glm(pres_abs2 ~ species*hemi, data=pmc_species_morphology.chisq.prculs_d,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.prculs_d.glm.aov <- Anova(pmc_species_morphology.chisq.prculs_d.glm, test="LR", type="III")
pmc_species_morphology.chisq.prculs_d.glm.aov


## prcus-p
### Run analysis
pmc_species_morphology.chisq.prcus_p <- pmc_species_morphology.chisq %>% subset(label == "prcus_p")
pmc_species_morphology.chisq.prcus_p.glm <- glm(pres_abs2 ~ species*hemi, data=pmc_species_morphology.chisq.prcus_p,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.prcus_p.glm.aov <- Anova(pmc_species_morphology.chisq.prcus_p.glm, test="LR", type="III")
pmc_species_morphology.chisq.prcus_p.glm.aov


## prcus-i
### Run analysis
pmc_species_morphology.chisq.prcus_i <- pmc_species_morphology.chisq %>% subset(label == "prcus_i")
pmc_species_morphology.chisq.prcus_i.glm <- glm(pres_abs2 ~ species*hemi, data=pmc_species_morphology.chisq.prcus_i,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.prcus_i.glm.aov <- Anova(pmc_species_morphology.chisq.prcus_i.glm, test="LR", type="III")
pmc_species_morphology.chisq.prcus_i.glm.aov


## prcus-a
### Run analysis
pmc_species_morphology.chisq.prcus_a <- pmc_species_morphology.chisq %>% subset(label == "prcus_a")
pmc_species_morphology.chisq.prcus_a.glm <- glm(pres_abs2 ~ species*hemi, data = pmc_species_morphology.chisq.prcus_a,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.prcus_a.glm.aov <- Anova(pmc_species_morphology.chisq.prcus_a.glm, test="LR", type="III")
pmc_species_morphology.chisq.prcus_a.glm.aov


## isms
### Run analysis
pmc_species_morphology.chisq.isms <- pmc_species_morphology.chisq %>% subset(label == "isms")
pmc_species_morphology.chisq.isms.glm <- glm(pres_abs2 ~ species*hemi, data = pmc_species_morphology.chisq.isms,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.isms.glm.aov <- Anova(pmc_species_morphology.chisq.isms.glm, test="LR", type="III")
pmc_species_morphology.chisq.isms.glm.aov


## prculs-v
### Run analysis
pmc_species_morphology.chisq.prculs_v <- pmc_species_morphology.chisq %>% subset(label == "prculs_v")
pmc_species_morphology.chisq.prculs_v.glm <- glm(pres_abs2 ~ species*hemi, data = pmc_species_morphology.chisq.prculs_v,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.prculs_v.glm.aov <- Anova(pmc_species_morphology.chisq.prculs_v.glm, test="LR", type="III")
pmc_species_morphology.chisq.prculs_v.glm.aov

## sspls-v
### Run analysis
pmc_species_morphology.chisq.sspls_v <- pmc_species_morphology.chisq %>% subset(label == "sspls_v")
pmc_species_morphology.chisq.sspls_v.glm <- glm(pres_abs2 ~ species*hemi, data = pmc_species_morphology.chisq.sspls_v,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.sspls_v.glm.aov <- Anova(pmc_species_morphology.chisq.sspls_v.glm, test="LR", type="III")
pmc_species_morphology.chisq.sspls_v.glm.aov

### Digest interaction
pmc_species_morphology.chisq.sspls_v.glm.i1 <- emmeans::emmeans(pmc_species_morphology.chisq.sspls_v.glm,'species','hemi')
emmeans::contrast(pmc_species_morphology.chisq.sspls_v.glm.i1, method='pairwise')


## sspls-d
### Run analysis
pmc_species_morphology.chisq.sspls_d <- pmc_species_morphology.chisq %>% subset(label == "sspls_d")
pmc_species_morphology.chisq.sspls_d.glm <- glm(pres_abs2 ~ species*hemi, data = pmc_species_morphology.chisq.sspls_d,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.sspls_d.glm.aov <- Anova(pmc_species_morphology.chisq.sspls_d.glm, test="LR", type="III")
pmc_species_morphology.chisq.sspls_d.glm.aov


## ifrms
### Run analysis
pmc_species_morphology.chisq.ifrms <- pmc_species_morphology.chisq %>% subset(label == "ifrms")
pmc_species_morphology.chisq.ifrms.glm <- glm(pres_abs2 ~ species*hemi, data = pmc_species_morphology.chisq.ifrms,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.ifrms.glm.aov <- Anova(pmc_species_morphology.chisq.ifrms.glm, test="LR", type="III")
pmc_species_morphology.chisq.ifrms.glm.aov


## icgs-p
### Run analysis
pmc_species_morphology.chisq.icgs_p <- pmc_species_morphology.chisq %>% subset(label == "icgs_p")
pmc_species_morphology.chisq.icgs_p.glm <- glm(pres_abs2 ~ species*hemi, data = pmc_species_morphology.chisq.icgs_p,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.icgs_p.glm.aov <- Anova(pmc_species_morphology.chisq.icgs_p.glm, test="LR", type="III")
pmc_species_morphology.chisq.icgs_p.glm.aov


## pmcgs (w)
### Run analysis
pmc_species_morphology.chisq.w <- pmc_species_morphology.chisq %>% subset(label == "w")
pmc_species_morphology.chisq.w.glm <- glm(pres_abs2 ~ species*hemi, data = pmc_species_morphology.chisq.w,
           family=binomial(link = "logit"))

### View summary
pmc_species_morphology.chisq.w.glm.aov <- Anova(pmc_species_morphology.chisq.w.glm, test="LR", type="III")
pmc_species_morphology.chisq.w.glm.aov


# Analysis 2.3

## Set up data for analyses
pmc_species_morphology.chisq.pcc.reg <- pmc_species_morphology.chisq %>% 
  subset(label %in% c("icgs_p", "ifrms", "isms",  "sspls_d", "sspls_v")) %>% 
  mutate(
    PCC_region = case_when(label %in% c("icgs_p", "ifrms", "sspls_d") ~ "dPCC",
                          label %in% c("sspls_v", "isms") ~ "vPCC")
  )

## Run analysis
pmc_species_morphology.chisq.pcc.reg.glm <- glm(pres_abs2 ~ species*hemi*PCC_region, data = pmc_species_morphology.chisq.pcc.reg,
           family=binomial(link = "logit"))

## View summary
pmc_species_morphology.chisq.pcc.reg.glm.aov <- Anova(pmc_species_morphology.chisq.pcc.reg.glm, test="LR", type="III")
pmc_species_morphology.chisq.pcc.reg.glm.aov

## Digest interaction
pmc_species_morphology.chisq.pcc.reg.glm.i1 <- emmeans::emmeans(pmc_species_morphology.chisq.pcc.reg.glm, 'species', 'PCC_region')
emmeans::contrast(pmc_species_morphology.chisq.pcc.reg.glm.i1, method='pairwise', adjust = "tukey")
```

## Plots (Figures 2c, 3, 4, 5)
```{r}
# Figure 2c

## Manipulate data for PrC plot
pmc_species_morphology.chisq.subject.prc.mean <- pmc_species_morphology.chisq.subject.prc %>% 
  group_by(species, hemi) %>% 
  summarise(mean = round(mean(n), 1))

pmc_species_morphology.chisq.subject.prc.n <- pmc_species_morphology.chisq.subject.prc %>% group_by(species, hemi, n) %>% summarise(n2 = n())

## Make PrC plot (Figure 2c, left)
pmc_species_morphology.chisq.subject.prc.plot <- pmc_species_morphology.chisq.subject.prc.n %>% 
  ggplot(aes(x = n, y = n2)) +
  
  geom_col(aes(fill = species), color = "black") +
  
  # geom_vline(xintercept = 2, color = "red") +
  # 
  # geom_vline(xintercept = 1.8, color = "darkred") +
  # 
  # geom_vline(xintercept = 4.4, color = "blue") +
  # 
  # geom_vline(xintercept = 4.4, color = "navy") +
  
  scale_fill_manual(breaks = c("chimpanzee", "humans"),
                     values = c("#cccccc", "#252525")) +
  
  facet_wrap(~ hemi, 
             nrow = 1, ncol = 2, 
             strip.position = "top"
             ) + 
  
  scale_x_continuous(n.breaks = 6) +
  scale_y_continuous(n.breaks = 5, limits = c(0,50)) +
  
  guides(fill = "none", alpha = "none") +
  
  plot_theme2

## View plot
pmc_species_morphology.chisq.subject.prc.plot 

## Save plot
# ggplot2::ggsave(filename = "Figure_2c_prc_incidence_count.png",
#                  plot = pmc_species_morphology.chisq.subject.prc.plot,
#                  device = "png",
#                  width = 8, # 12
#                  height = 5,
#                  units = "in",
#                  dpi = "retina") 


## Manipulate data for PCC plot
pmc_species_morphology.chisq.subject.pcc.mean <- pmc_species_morphology.chisq.subject.pcc %>% 
  group_by(species, hemi) %>% 
  summarise(mean = round(mean(n), 1))
pmc_species_morphology.chisq.subject.pcc.mean

pmc_species_morphology.chisq.subject.pcc.n <- pmc_species_morphology.chisq.subject.pcc %>% group_by(species, hemi, n) %>% summarise(n2 = n())

## Make PCC plot (Figure 2c, right)
pmc_species_morphology.chisq.subject.pcc.plot <- pmc_species_morphology.chisq.subject.pcc.n %>% 
  ggplot(aes(x = n, y = n2)) +
  
  geom_col(aes(fill = species), color = "black") +
  
  # geom_vline(xintercept = 1.7, color = 'red') +
  # 
  # geom_vline(xintercept = 1.8, color = 'darkred') +
  # 
  # geom_vline(xintercept = 2.9, color = 'blue') +
  # 
  # geom_vline(xintercept = 2.8, color = 'navy') +
  
  scale_fill_manual(breaks = c("chimpanzee", "humans"),
                     values = c("#cccccc", "#252525")) +
  
  facet_wrap(~ hemi, 
             nrow = 1, ncol = 2, 
             strip.position = "top"
             ) + 
  
  scale_x_continuous(n.breaks = 6) +
  scale_y_continuous(n.breaks = 5) +
  
  guides(fill = "none", alpha = "none") +
  
  plot_theme2
 
## View plot 
pmc_species_morphology.chisq.subject.pcc.plot

## Save plot
# ggplot2::ggsave(filename = "Figure_2c_pcc_incidence_count.png",
#                  plot = pmc_species_morphology.chisq.subject.pcc.plot,
#                  device = "png",
#                  width = 8, # 12
#                  height = 5,
#                  units = "in",
#                  dpi = "retina") 


# Figures 3-5

## Load data
pmc_incidence_table_figure <- read.csv("pmc_pres_abs.csv")

## Manipulations for plot
pmc_incidence_table_figure.fig <- pmc_incidence_table_figure %>% subset(pres == "present")
pmc_incidence_table_figure.fig$hemi <- gsub("lh", "Left Hemisphere", pmc_incidence_table_figure.fig$hemi)
pmc_incidence_table_figure.fig$hemi <- gsub("rh", "Right Hemisphere", pmc_incidence_table_figure.fig$hemi)
pmc_incidence_table_figure.fig$sulcus <- gsub("w", "pmcgs", pmc_incidence_table_figure.fig$sulcus)

pmc_incidence_table_figure.fig$sulcus <- factor(pmc_incidence_table_figure.fig$sulcus, levels = c("pos", "prculs-d", "prculs-v",
                                                                                                  "prcus-p", "prcus-i", "prcus-a",
                                                                                                  "isms", "sspls-v", "sspls-d",
                                                                                                  "ifrms", "icgs-p", "spls", "mcgs", "pmcgs"
                                                                                          ))

pmc_incidence_table_figure.fig <- pmc_incidence_table_figure.fig %>% mutate(is_prcus_i = case_when(
                                sulcus == "prcus-i" ~ "is_prcus_i",
                                sulcus %in% c("mcgs", "pmcgs", "spls", "pos", 
                                              "isms", "sspls-v", "sspls-d", "ifrms", "icgs-p", "prcus-a", 
                                              "prcus-p", "prculs-v", "prculs-d") ~ "not_prcus_i")) %>%
                                mutate(
                                is_sspls_v_or_isms = case_when(
                                sulcus %in% c("isms", "sspls-v") ~ "is_sspls_v_or_isms",
                                sulcus %in% c("mcgs", "pmcgs", "spls", "pos", 
                                              "sspls-d", "ifrms", "icgs-p", "prcus-a", "prcus-i",
                                              "prcus-p", "prculs-v", "prculs-d") ~ "not_sspls_v_or_isms")) %>%
  
                                mutate(
                                border_new = case_when(
                                sulcus %in% c( "pmcgs") ~ "new_border",
                                sulcus %in% c("mcgs",  "isms", "sspls-v", "spls", "pos", 
                                              "sspls-d", "ifrms", "icgs-p", "prcus-a", "prcus-i",
                                              "prcus-p", "prculs-v", "prculs-d") ~ "not_new_border")) %>%
  
                                mutate(
                                prc_new = case_when(
                                sulcus %in% c( "prculs-v" ) ~ "new_prc",
                                sulcus %in% c( "mcgs", "pmcgs", "isms", "sspls-v", "spls", "pos", 
                                              "sspls-d", "ifrms", "icgs-p", "prcus-a", "prcus-i",
                                              "prcus-p", "prculs-d") ~ "not_new_prc")) %>%
                                
                                mutate(
                                pcc_new = case_when(
                                sulcus %in% c( "isms", "sspls-v" ) ~ "new_pcc",
                                sulcus %in% c( "mcgs", "pmcgs", "isms", "sspls-v", "spls", "pos", 
                                              "sspls-d", "ifrms", "icgs-p", "prcus-a", "prcus-i",
                                              "prcus-p", "prculs-v", "prculs-d") ~ "not_new_pcc"))
                            

## Make plot
### Comment in and out region of interest: Border, PrC, PCC for plot of interest
pmc_incidence.figure.prc <- pmc_incidence_table_figure.fig %>% 
  
  # For border
  #subset(sulcus %in% c("mcgs", "pmcgs", "spls", "pos")) %>%
  
  # For PrC
  subset(sulcus %in% c("prcus-a", "prcus-i", "prcus-p", "prculs-v", "prculs-d")) %>%
  
  # For PCC
  #subset(sulcus %in% c("isms", "sspls-v", "sspls-d", "ifrms", "icgs-p")) %>%
  
  ggplot(aes(x = sulcus, y = pct)) + 
  
  geom_col(aes(fill = sulcus, alpha = species),
           color = "black",
           position = "dodge"
           ) +
  
  facet_wrap(~ hemi, 
             nrow = 1, ncol = 2, 
             strip.position = "top"
             ) + 
  
  # For border
  # scale_fill_manual(breaks = c("pos", "spls", "mcgs", "pmcgs"),
  #                  values = c("#6850A0", "#518DF0", "#17314B", "#B5CFFF")) +
  # 

  # For PrC
  scale_fill_manual(breaks = c("prculs-d", "prculs-v", "prcus-p", "prcus-i", "prcus-a"),
                    values = c("#858CC2", "#9AA78C", "#66B9B8", "#2F5952", "#45787F")
                    ) +

  # For PCC
  # scale_fill_manual(breaks = c("isms", "sspls-v", "sspls-d", "ifrms", "icgs-p"),
  #                   values = c("#DC5075", "#DEC65B", "#9D6642", "#992F27", "#E98B32"),
  #                   ) +
  
  scale_alpha_manual(breaks = c("chimpanzee", "humans"),
                     values = c(0.5, 1)) +
  
  scale_y_continuous(name = "% of hemispheres", seq(0,100,20), limits = c(0,100)) +
  
  guides(fill = "none", alpha = "none", pattern = "none") +
  plot_theme2

## View plot
pmc_incidence.figure.prc

## Save plot
# ggplot2::ggsave(filename = "Figure_4_pmc_incidence_prc.png",
#                  plot = pmc_incidence.figure.prc,
#                  device = "png",
#                  width = 12, # 10
#                  height = 5,
#                  units = "in",
#                  dpi = "retina")
```


# Analysis 3: Relating incidence of PMC sulci in chimpanzees 
1. We ran binomial logistic regression GLMs relating the presence of each variable PMC sulcus [0 (absent), 1 (present)] with the other sulci as factors, while also including an interaction with hemisphere for each sulcus. 
* Iteratively dropped the sulcus that was the dependent variable as a factor from the next model to account for relationships already analyzed.

## Statistics
```{r}
# Prepare left hemisphere data
pmc_chimp_morphology.ifrms.lh <- chimp_morphology %>% 
  subset(label == "ifrms" & hemi == "lh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(ifrms_pres = pres_abs)

pmc_chimp_morphology.isms.lh <- chimp_morphology %>% 
  subset(label == "isms" & hemi == "lh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(isms_pres = pres_abs)

pmc_chimp_morphology.sspls_v.lh <- chimp_morphology %>% 
  subset(label == "sspls_v" & hemi == "lh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(sspls_v_pres = pres_abs)

pmc_chimp_morphology.prcus_i.lh <- chimp_morphology %>% 
  subset(label == "prcus_i" & hemi == "lh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(prcus_i_pres = pres_abs)

pmc_chimp_morphology.prcus_a.lh <- chimp_morphology %>% 
  subset(label == "prcus_a" & hemi == "lh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(prcus_a_pres = pres_abs)

pmc_chimp_morphology.prcus_p.lh <- chimp_morphology %>% 
  subset(label == "prcus_p" & hemi == "lh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(prcus_p_pres = pres_abs)

pmc_chimp_morphology.icgs_p.lh <- chimp_morphology %>% 
  subset(label == "icgs_p" & hemi == "lh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(icgs_p_pres = pres_abs)

pmc_chimp_morphology.pres_abs.lh <- merge(pmc_chimp_morphology.ifrms.lh, pmc_chimp_morphology.isms.lh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.lh <- merge(pmc_chimp_morphology.pres_abs.lh, pmc_chimp_morphology.sspls_v.lh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.lh <- merge(pmc_chimp_morphology.pres_abs.lh, pmc_chimp_morphology.prcus_i.lh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.lh <- merge(pmc_chimp_morphology.pres_abs.lh, pmc_chimp_morphology.prcus_a.lh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.lh <- merge(pmc_chimp_morphology.pres_abs.lh, pmc_chimp_morphology.prcus_p.lh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.lh <- merge(pmc_chimp_morphology.pres_abs.lh, pmc_chimp_morphology.icgs_p.lh, by = c("sub", "hemi"))

pmc_chimp_morphology.pres_abs.lh.analysis <- pmc_chimp_morphology.pres_abs.lh %>% select(c(1, 3:9))
pmc_chimp_morphology.pres_abs.lh.analysis <- pmc_chimp_morphology.pres_abs.lh.analysis %>% 
                                                                            mutate(ifrms_pres2 = case_when(ifrms_pres == "present" ~ 1,
                                                                                                           ifrms_pres == "absent" ~ 0),
                                                                                   isms_pres2 = case_when(isms_pres == "present" ~ 1,
                                                                                                          isms_pres == "absent" ~ 0),
                                                                                   sspls_v_pres2 = case_when(sspls_v_pres == "present" ~ 1,
                                                                                                             sspls_v_pres == "absent" ~ 0),
                                                                                   prcus_i_pres2 = case_when(prcus_i_pres == "present" ~ 1,
                                                                                                             prcus_i_pres == "absent" ~ 0),
                                                                                   prcus_a_pres2 = case_when(prcus_a_pres == "present" ~ 1,
                                                                                                             prcus_a_pres == "absent" ~ 0),
                                                                                   prcus_p_pres2 = case_when(prcus_p_pres == "present" ~ 1,
                                                                                                             prcus_p_pres == "absent" ~ 0),
                                                                                   icgs_p_pres2 = case_when(icgs_p_pres == "present" ~ 1,
                                                                                                            icgs_p_pres == "absent" ~ 0),
                                                                                   )


# Prepare right hemisphere data
pmc_chimp_morphology.ifrms.rh <- chimp_morphology %>% 
  subset(label == "ifrms" & hemi == "rh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(ifrms_pres = pres_abs)

pmc_chimp_morphology.isms.rh <- chimp_morphology %>% 
  subset(label == "isms" & hemi == "rh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(isms_pres = pres_abs)

pmc_chimp_morphology.sspls_v.rh <- chimp_morphology %>% 
  subset(label == "sspls_v" & hemi == "rh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(sspls_v_pres = pres_abs)

pmc_chimp_morphology.prcus_i.rh <- chimp_morphology %>% 
  subset(label == "prcus_i" & hemi == "rh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(prcus_i_pres = pres_abs)

pmc_chimp_morphology.prcus_a.rh <- chimp_morphology %>% 
  subset(label == "prcus_a" & hemi == "rh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(prcus_a_pres = pres_abs)

pmc_chimp_morphology.prcus_p.rh <- chimp_morphology %>% 
  subset(label == "prcus_p" & hemi == "rh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(prcus_p_pres = pres_abs)

pmc_chimp_morphology.icgs_p.rh <- chimp_morphology %>% 
  subset(label == "icgs_p" & hemi == "rh") %>% 
  select(c("sub", "hemi", "pres_abs")) %>% 
  rename(icgs_p_pres = pres_abs)

pmc_chimp_morphology.pres_abs.rh <- merge(pmc_chimp_morphology.ifrms.rh, pmc_chimp_morphology.isms.rh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.rh <- merge(pmc_chimp_morphology.pres_abs.rh, pmc_chimp_morphology.sspls_v.rh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.rh <- merge(pmc_chimp_morphology.pres_abs.rh, pmc_chimp_morphology.prcus_i.rh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.rh <- merge(pmc_chimp_morphology.pres_abs.rh, pmc_chimp_morphology.prcus_a.rh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.rh <- merge(pmc_chimp_morphology.pres_abs.rh, pmc_chimp_morphology.prcus_p.rh, by = c("sub", "hemi"))
pmc_chimp_morphology.pres_abs.rh <- merge(pmc_chimp_morphology.pres_abs.rh, pmc_chimp_morphology.icgs_p.rh, by = c("sub", "hemi"))

pmc_chimp_morphology.pres_abs.rh.analysis <- pmc_chimp_morphology.pres_abs.rh %>% select(c(1, 3:9))
pmc_chimp_morphology.pres_abs.rh.analysis <- pmc_chimp_morphology.pres_abs.rh.analysis %>% 
                                                                            mutate(ifrms_pres2 = case_when(ifrms_pres == "present" ~ 1,
                                                                                                           ifrms_pres == "absent" ~ 0),
                                                                                   isms_pres2 = case_when(isms_pres == "present" ~ 1,
                                                                                                          isms_pres == "absent" ~ 0),
                                                                                   sspls_v_pres2 = case_when(sspls_v_pres == "present" ~ 1,
                                                                                                             sspls_v_pres == "absent" ~ 0),
                                                                                   prcus_i_pres2 = case_when(prcus_i_pres == "present" ~ 1,
                                                                                                             prcus_i_pres == "absent" ~ 0),
                                                                                   prcus_a_pres2 = case_when(prcus_a_pres == "present" ~ 1,
                                                                                                             prcus_a_pres == "absent" ~ 0),
                                                                                   prcus_p_pres2 = case_when(prcus_p_pres == "present" ~ 1,
                                                                                                             prcus_p_pres == "absent" ~ 0),
                                                                                   icgs_p_pres2 = case_when(icgs_p_pres == "present" ~ 1,
                                                                                                            icgs_p_pres == "absent" ~ 0),
                                                                                   )

# Merge data from left and right hemispheres
pmc_chimp_morphology.pres_abs.lh.analysis$hemi <- "lh"
pmc_chimp_morphology.pres_abs.rh.analysis$hemi <- "rh"
pmc_chimp_morphology.pres_abs.analysis <- rbind(pmc_chimp_morphology.pres_abs.lh.analysis, pmc_chimp_morphology.pres_abs.rh.analysis)

# Make data a factor
pmc_chimp_morphology.pres_abs.analysis$ifrms_pres2 <- as.factor(pmc_chimp_morphology.pres_abs.analysis$ifrms_pres2)
pmc_chimp_morphology.pres_abs.analysis$isms_pres2 <- as.factor(pmc_chimp_morphology.pres_abs.analysis$isms_pres2)
pmc_chimp_morphology.pres_abs.analysis$sspls_v_pres2 <- as.factor(pmc_chimp_morphology.pres_abs.analysis$sspls_v_pres2)
pmc_chimp_morphology.pres_abs.analysis$prcus_i_pres2 <- as.factor(pmc_chimp_morphology.pres_abs.analysis$prcus_i_pres2)
pmc_chimp_morphology.pres_abs.analysis$prcus_a_pres2 <- as.factor(pmc_chimp_morphology.pres_abs.analysis$prcus_a_pres2)
pmc_chimp_morphology.pres_abs.analysis$prcus_p_pres2 <- as.factor(pmc_chimp_morphology.pres_abs.analysis$prcus_p_pres2)
pmc_chimp_morphology.pres_abs.analysis$icgs_p_pres2 <- as.factor(pmc_chimp_morphology.pres_abs.analysis$icgs_p_pres2)


# Analysis 3

## Model 1: Sspls-v ~ isms, ifrms, icgs_p 
### Run analysis
fit_sspls_v <- glm(sspls_v_pres2 ~ isms_pres2 + isms_pres2:hemi + 
                                   ifrms_pres2 + ifrms_pres2:hemi + 
                                   prcus_i_pres2 + prcus_i_pres2:hemi +
                                   prcus_p_pres2 + prcus_p_pres2:hemi,
                   data=pmc_chimp_morphology.pres_abs.analysis,
           family=binomial())

### View summary
Anova(fit_sspls_v, test="LR", type="III")

### Get odds ratio for effect
fit_sspls_v %>%
  broom::tidy(exponentiate = T, conf.int = T) %>%
  mutate_if(is.numeric, ~ round(.,3)) 
100*(4.47-1)


## Model 2: Isms, dropped sspls_v from prior analysis
### Run analysis
fit_isms <- glm(isms_pres2 ~ ifrms_pres2 + ifrms_pres2:hemi +
                               prcus_i_pres2 + prcus_i_pres2:hemi +
                               prcus_p_pres2 + prcus_p_pres2:hemi,
                 data=pmc_chimp_morphology.pres_abs.analysis,
           family=binomial())

### View summary
Anova(fit_isms, test="LR", type="III")


## Model 3: Ifrms, dropped sspls-v and isms from prior analyses 
### Run analysis
fit_ifrms <- glm(ifrms_pres2 ~ prcus_i_pres2 + prcus_i_pres2:hemi +
                                 prcus_p_pres2 + prcus_p_pres2:hemi, 
                   data=pmc_chimp_morphology.pres_abs.analysis,
           family=binomial())

### View summary
Anova(fit_ifrms, test="LR", type="III")


## Model 4: Prcus-i, dropping sspls-v, isms, and ifrms from prior analyses
### Run analysis
fit_prcus_i <- glm(prcus_i_pres2 ~ prcus_p_pres2 + prcus_p_pres2:hemi, 
                   data=pmc_chimp_morphology.pres_abs.analysis,
           family=binomial())

### View summary
Anova(fit_prcus_i, test="LR", type="III")
```

## Plot (Figure 6)
```{r}
# Figure 6

# Set palette for plot
my_palette2 <- brewer.pal(2,"Greens")

# Manipulate data for plot
sspls_v_pred.tab <- pmc_chimp_morphology.pres_abs.analysis %>% group_by(sspls_v_pres, isms_pres) %>% summarise(n = n(),
                                                                                                               pct = n/120*100)

# Make plot
sspls_v_pred.plot <- sspls_v_pred.tab %>% 
  ggplot(aes(x = sspls_v_pres, y = pct, fill = isms_pres)) + 
  geom_col(color = 'black') +
  scale_y_continuous(name = "% of hemispheres", n.breaks = 6) +
  scale_fill_manual(breaks = c("present", "absent"),
                    values = c("#252525", "#cccccc")
                    ) +
  guides(fill = "none") +
  plot_theme2
  
# View plot
sspls_v_pred.plot

# Save plot
# ggplot2::ggsave(filename = "Figure_6_sspls_v_isms_incidence.png",
#                  plot = sspls_v_pred.plot,
#                  device = "png",
#                  width = 4,
#                  height = 5,
#                  units = "in",
#                  dpi = "retina")
```


# Analysis 4: Comparing morphology of PMC sulci between chimpanzees and humans
1. To assess whether the depth and surface area of PMC sulci differed between chimpanzees and humans, for both morphological features, we ran a LME with predictors of sulcus, hemisphere, and species, as well as their interaction. 
* Species, hemisphere, and sulcus were considered fixed effects. 
* Sulcus was nested within the hemisphere which was nested within subjects. 

## Statistics: Depth
```{r}
# Analysis 4.1: Depth

## Select necessary columns and sulci with incidence > 50%
pmc_species_morphology.depth <- pmc_species_morphology %>% subset(label %in% c("pos", "prculs_d", "prcus_i", "spls", "mcgs", "w", "sspls_v", "isms"))

## Normalize depth
pmc_species_morphology.depth <- merge(pmc_species_morphology.depth, insula_depth, by = c("sub", "hemi", "species"))
pmc_species_morphology.depth <- pmc_species_morphology.depth %>% mutate(normalized_depth = sulcal_depth_mm/insula_depth)

## Remove NA values
pmc_species_morphology.depth.analysis <- pmc_species_morphology.depth %>% drop_na()

## Run analysis
model_depth_species <- lme(normalized_depth ~ hemi*label*species, random = (~1|sub/hemi/label), data = pmc_species_morphology.depth.analysis)

## View summary
model_depth_species.aov <- anova(model_depth_species)
model_depth_species.aov

## Effect size
effectsize::eta_squared(model_depth_species.aov)

## Digest main effects and interactions
model_depth_species.m1 <- emmeans::emmeans(model_depth_species, ~ species)
emmeans::contrast(model_depth_species.m1, method='pairwise')

model_depth_species.i1 <- emmeans::emmeans(model_depth_species, ~ species | hemi)
emmeans::contrast(model_depth_species.i1, method='pairwise')

model_depth_species.i2 <- emmeans::emmeans(model_depth_species, ~ species | label)
emmeans::contrast(model_depth_species.i2, method='pairwise')

model_depth_species.i3 <- emmeans::emmeans(model_depth_species, ~ species | label | hemi)
emmeans::contrast(model_depth_species.i3, method='pairwise')
```

## Function for morphology plots: Split violin
```{r}
GeomSplitViolin <- ggproto("GeomSplitViolin", GeomViolin, 
                       draw_group = function(self, data, ..., draw_quantiles = NULL) {
    data <- transform(data, xminv = x - violinwidth * (x - xmin), xmaxv = x + violinwidth * (xmax - x))
   grp <- data[1, "group"]
   newdata <- plyr::arrange(transform(data, x = if (grp %% 2 == 1) xminv else xmaxv), if (grp %% 2 == 1) y else -y)
   newdata <- rbind(newdata[1, ], newdata, newdata[nrow(newdata), ], newdata[1, ])
   newdata[c(1, nrow(newdata) - 1, nrow(newdata)), "x"] <- round(newdata[1, "x"])
   if (length(draw_quantiles) > 0 & !scales::zero_range(range(data$y))) {
     stopifnot(all(draw_quantiles >= 0), all(draw_quantiles <=
       1))
     quantiles <- ggplot2:::create_quantile_segment_frame(data, draw_quantiles)
     aesthetics <- data[rep(1, nrow(quantiles)), setdiff(names(data), c("x", "y")), drop = FALSE]
     aesthetics$alpha <- rep(1, nrow(quantiles))
     both <- cbind(quantiles, aesthetics)
     quantile_grob <- GeomPath$draw_panel(both, ...)
     ggplot2:::ggname("geom_split_violin", grid::grobTree(GeomPolygon$draw_panel(newdata, ...), quantile_grob))
   }
   else {
     ggplot2:::ggname("geom_split_violin", GeomPolygon$draw_panel(newdata, ...))
   }
 })
 geom_split_violin <- function(mapping = NULL, data = NULL, stat = "ydensity", position = "identity", ..., 
                               draw_quantiles = NULL, trim = TRUE, scale = "area", na.rm = FALSE, 
                               show.legend = NA, inherit.aes = TRUE) {
   layer(data = data, mapping = mapping, stat = stat, geom = GeomSplitViolin, 
         position = position, show.legend = show.legend, inherit.aes = inherit.aes, 
         params = list(trim = trim, scale = scale, draw_quantiles = draw_quantiles, na.rm = na.rm, ...))
 }
```

## Plots: Depth (Figure 7a)
```{r}
# Figure 7a 

# Manipulation for plot
pmc_species_morphology.depth.analysis$hemi <- gsub("lh", "left", pmc_species_morphology.depth.analysis$hemi)
pmc_species_morphology.depth.analysis$hemi <- gsub("rh", "right", pmc_species_morphology.depth.analysis$hemi)
pmc_species_morphology.depth.analysis$label <- gsub("w", "pmcgs", pmc_species_morphology.depth.analysis$label)

pmc_species_morphology.depth.analysis$label <- gsub("prcus_i", "prcus-i", pmc_species_morphology.depth.analysis$label)
pmc_species_morphology.depth.analysis$label <- gsub("prculs_d", "prculs-d", pmc_species_morphology.depth.analysis$label)
pmc_species_morphology.depth.analysis$label <- gsub("sspls_v", "sspls-v", pmc_species_morphology.depth.analysis$label)


pmc_species_morphology.depth.analysis$label <- factor(pmc_species_morphology.depth.analysis$label, levels = c(
   "mcgs", "pmcgs", "spls", "prcus-i", "prculs-d", "pos", "isms", "sspls-v"
   ))

# Make plot 
pmc_species_morphology.depth.analysis.plot <- pmc_species_morphology.depth.analysis %>% 
  ggplot(aes(x = label, y = normalized_depth*100, alpha = species, fill = label)) + 
  
  geom_split_violin(width = 1.5) + 
  
  geom_boxplot(
               position = position_dodge(width =.14),
               outlier.size = 0, outlier.alpha = 0, 
               width = .13) +
  
  scale_y_continuous(name = "sulcal depth (% max)") +
  
  facet_wrap(~ hemi, nrow = 2) + 
  
  scale_fill_manual(breaks = c("mcgs", "pmcgs", "spls", "prcus-i", "prculs-d", "pos", "isms", "sspls-v"),
                    values = c("#17314B", "#B5CFFF", "#518DF0", 
                               "#2F5952", "#858CC2", "#6850A0", "#DC5075", "#DEC65B"),
                    ) +
  
  scale_alpha_manual(breaks = c("chimpanzee", "human"),
                     values = c(0.3, 0.65)) +
  
  guides(fill = "none", alpha = "none") +

  plot_theme2

# View plot 
pmc_species_morphology.depth.analysis.plot

# Save plot
# ggplot2::ggsave(filename = "Figure_7a_pmc_species_morphology_depth.png",
#                  plot = pmc_species_morphology.depth.analysis.plot,
#                  device = "png",
#                  width = 8,
#                  height = 5,
#                  units = "in",
#                  dpi = "retina")
```

## Statistics: Surface area
```{r}
# Analysis 4.2: Surface area

## Select necessary columns and sulci with incidence > 50%
pmc_species_morphology.SA <- pmc_species_morphology %>% subset(label %in% c("pos", "prculs_d", "prcus_i", "spls", "mcgs", "w", "sspls_v", "isms"))

## Normalize surface area
pmc_species_morphology.SA <- merge(pmc_species_morphology.SA, lobe_morphology, by = c("sub", "hemi", "species"))
pmc_species_morphology.SA <- pmc_species_morphology.SA %>% mutate(normalized_SA = total_surface_area_.mm.2./cortex_sa)

## Remove NA values
pmc_species_morphology.SA <- pmc_species_morphology.SA %>% drop_na()

## Run analysis
model_SA_species <- lme(normalized_SA ~ hemi*label*species, random = (~1|sub/hemi/label), data = pmc_species_morphology.SA)

## View summary
model_SA_species.aov <- anova(model_SA_species)
model_SA_species.aov

## Effect size
effectsize::eta_squared(model_SA_species.aov)

## Digest main effects and interactions
model_sa_species.m1 <- emmeans::emmeans(model_SA_species, ~ species)
emmeans::contrast(model_sa_species.m1, method='pairwise', adjust='tukey')

model_SA_species.i1 <- emmeans::emmeans(model_SA_species, ~ species | hemi)
emmeans::contrast(model_SA_species.i1, method='pairwise')

model_SA_species.i2 <- emmeans::emmeans(model_SA_species, ~ species | label)
emmeans::contrast(model_SA_species.i2, method='pairwise')

model_SA_species.i3 <- emmeans::emmeans(model_SA_species, ~ species | label | hemi)
emmeans::contrast(model_SA_species.i3, method='pairwise')
```

## Plots: Surface Area (Figure 7b)
```{r}
# Figure 7b

# Manipulation for plot
pmc_species_morphology.SA$hemi <- gsub("lh", "left", pmc_species_morphology.SA$hemi)
pmc_species_morphology.SA$hemi <- gsub("rh", "right", pmc_species_morphology.SA$hemi)
pmc_species_morphology.SA$label <- gsub("w", "pmcgs", pmc_species_morphology.SA$label)

pmc_species_morphology.SA$label <- gsub("prcus_i", "prcus-i", pmc_species_morphology.SA$label)
pmc_species_morphology.SA$label <- gsub("prculs_d", "prculs-d", pmc_species_morphology.SA$label)
pmc_species_morphology.SA$label <- gsub("sspls_v", "sspls-v", pmc_species_morphology.SA$label)

pmc_species_morphology.SA$label <- factor(pmc_species_morphology.SA$label, levels = c(
   "mcgs", "pmcgs", "spls", "prcus-i", "prculs-d", "pos", "isms", "sspls-v"
   ))

pmc_species_morphology.SA$hemi <- gsub("lh", "Left Hemisphere", pmc_species_morphology.SA$hemi)
pmc_species_morphology.SA$hemi <- gsub("rh", "Right Hemisphere", pmc_species_morphology.SA$hemi)

# Make plot
pmc_species_morphology.SA.plot <- pmc_species_morphology.SA %>% 
  
  ggplot(aes(x = label, y = normalized_SA*100, alpha = species, fill = label)) + 
  
  geom_split_violin(width = 1.2) + 
  
  geom_boxplot(
               position = position_dodge(width =.14),
               outlier.size = 0, outlier.alpha = 0, 
               width = .13) +
  
  scale_y_continuous(name = "surface area (% cortex)") +
  
  facet_wrap(~ hemi, nrow = 2) + 
  
  scale_fill_manual(breaks = c("mcgs", "pmcgs", "spls", "prcus-i", "prculs-d", "pos", "isms", "sspls-v"),
                    values = c("#17314B", "#B5CFFF", "#518DF0", 
                               "#2F5952", "#858CC2", "#6850A0", "#DC5075", "#DEC65B"),
                    ) +
  
  scale_alpha_manual(breaks = c("chimpanzee", "human"),
                     values = c(0.3, 0.65)) +
  guides(fill = "none", alpha = "none") +
  
  plot_theme2

# View plot
pmc_species_morphology.SA.plot

# Save plot
# ggplot2::ggsave(filename = "Figure_7b_pmc_species_morphology_surface_area.png",
#                  plot = pmc_species_morphology.SA.plot,
#                  device = "png",
#                  width = 8,
#                  height = 5,
#                  units = "in",
#                  dpi = "retina")
```


# Analysis 5: Comparing incidence of MCGS patterns in chimpanzees versus humans
1. We quantitatively determined whether the incidence rates of the four MCGS types differed by 1) species, and between 2) hemispheres for each species, with chi-squared (χ2) tests. 

## Statistics
```{r}
# Analysis 5.1

## Make tables for analyses
mcgs_patterns.lh <- mcgs_patterns %>% subset(hemi == "lh")
mcgs_patterns.rh <- mcgs_patterns %>% subset(hemi == "rh")

mcgs_table_lh <- table(mcgs_patterns.lh$species, mcgs_patterns.lh$mcgs_pattern)
mcgs_table_lh <- as.data.frame(mcgs_table_lh)
mcgs_table_lh$hemi <- "lh"

mcgs_table_rh <- table(mcgs_patterns.rh$species, mcgs_patterns.rh$mcgs_pattern)
mcgs_table_rh <- as.data.frame(mcgs_table_rh)
mcgs_table_rh$hemi <- "rh"


## Analyses 5.1: Species
### Run analysis left hemisphere
chi_species_lh <- xtabs(Freq~Var1+Var2, data = mcgs_table_lh)

### View Summary
chisq.test(chi_species_lh)

### Digest effect
chisq.multcomp(chi_species_lh)

### Run analysis right hemisphere
chi_species_rh <- xtabs(Freq~Var1+Var2, data = mcgs_table_rh)

### View summary
chisq.test(chi_species_rh) 

### Digest effect
chisq.multcomp(chi_species_rh) 


## Analyses 5.2: Hemisphere
### Run analysis human
mcgs_patterns_human <- mcgs_patterns %>% subset(species == "human")
mcgs_patterns_human <- table(mcgs_patterns_human$hemi, mcgs_patterns_human$mcgs_pattern)
mcgs_table_human <- as.data.frame(mcgs_patterns_human)
mcgs_table_human <- xtabs(Freq~Var1+Var2, data = mcgs_table_human)
chi_species_human <- chisq.test(mcgs_table_human) 

### View summary
chi_species_human


### Run analysis chimpanzee
mcgs_patterns_chimp <- mcgs_patterns %>% subset(species == "chimpanzee")
mcgs_table_chimp <- table(mcgs_patterns_chimp$hemi, mcgs_patterns_chimp$mcgs_pattern)
mcgs_table_chimp <- as.data.frame(mcgs_table_chimp)
mcgs_table_chimp <- xtabs(Freq~Var1+Var2, data = mcgs_table_chimp)
chi_species_chimpanzee <- chisq.test(mcgs_table_chimp)

### View summary
chi_species_chimpanzee
```

## Plot (Figure 8c)
```{r}
# Figure 8c
# Load data
mcgs_incidence_table_figure <- read.csv("mcgs_figure.csv")

# Manipulations for plot
mcgs_incidence_table_figure$sulcus <- factor(mcgs_incidence_table_figure$sulcus, 
                                  levels = c("single_cih", "single", "Y", "trident") 
                                  )
mcgs_incidence_table_figure$hemi <- gsub("lh", "Left Hemisphere", mcgs_incidence_table_figure$hemi)
mcgs_incidence_table_figure$hemi <- gsub("rh", "Right Hemisphere", mcgs_incidence_table_figure$hemi)

mcgs_incidence_table_figure$sulcus <- gsub("single", "Type I", mcgs_incidence_table_figure$sulcus)
mcgs_incidence_table_figure$sulcus <- gsub("Y", "Type II", mcgs_incidence_table_figure$sulcus)
mcgs_incidence_table_figure$sulcus <- gsub("Type I_cih", "Type III", mcgs_incidence_table_figure$sulcus)
mcgs_incidence_table_figure$sulcus <- gsub("trident", "Type IV", mcgs_incidence_table_figure$sulcus)

# Set palette for plot
mypalette3 <- brewer.pal(4, "Blues")

# Make plot
mcgs_incidence_table_figure.plot <- mcgs_incidence_table_figure %>% ggplot(aes(x = species, 
                          y = pct, 
                          fill = sulcus)) + 
  geom_col(color = "black") + 
  facet_wrap(~ hemi, 
             nrow = 2, 
             strip.position = "top") + 
  scale_fill_manual(breaks = c("Type I", "Type II", "Type III", "Type IV"),
                     values = mypalette3
                     ) +
  scale_y_continuous(name = "% of hemispheres", seq(0,100,20), limits = c(0,100)) +
  plot_theme2 +
  guides(fill = "none")

# View plot
mcgs_incidence_table_figure.plot

# Save plot
# ggplot2::ggsave(filename = "Figure_8c_mcgs_incidence.png",
#                  plot = mcgs_incidence_table_figure.plot,
#                  device = "png",
#                  width = 3,
#                  height = 5,
#                  units = "in",
#                  dpi = "retina")
```


# Analysis 6: Comparing morphology of MCGS components between chimpanzees from humans? 
1. We repeated the pmc morphological analysis, exchanging the factor of PMC sulci for the mcgs branch (main branch, dorsal branch, side branch). 

## Statistics: Depth
```{r}
# Analysis 6.1

# Select necessary columns and sulci with incidence
mcgs_species_morphology.depth <- mcgs_species_morphology %>% 
  select(c("sub", "species", "hemi", "label", "sulcal_depth_mm"))

# Normalize depth
mcgs_species_morphology.depth <- merge(mcgs_species_morphology.depth, insula_depth, by = c("sub", "hemi", "species"))
mcgs_species_morphology.depth <- mcgs_species_morphology.depth %>% mutate(normalized_depth = sulcal_depth_mm/insula_depth)

# Remove NA values
mcgs_species_morphology.depth <- mcgs_species_morphology.depth %>% drop_na()

# Run analysis
model_mcgs_depth_species <- lme(normalized_depth ~ hemi*label*species, random = (~1|sub/hemi/label), data = mcgs_species_morphology.depth)

# View summary
model_mcgs_depth_species.aov <- anova(model_mcgs_depth_species)
model_mcgs_depth_species.aov

# Effect size
effectsize::eta_squared(model_mcgs_depth_species.aov)

# Digest main effects and interactions
model_mcgs_depth_species.m <- emmeans::emmeans(model_mcgs_depth_species, ~ hemi)
emmeans::contrast(model_mcgs_depth_species.m, method='pairwise')

model_mcgs_depth_species.m1 <- emmeans::emmeans(model_mcgs_depth_species, ~ label)
emmeans::contrast(model_mcgs_depth_species.m1, method='pairwise')

model_mcgs_depth_species.m2 <- emmeans::emmeans(model_mcgs_depth_species, ~ species)
emmeans::contrast(model_mcgs_depth_species.m2, method='pairwise')

model_mcgs_depth_species.i1 <- emmeans::emmeans(model_mcgs_depth_species, ~ species | label)
emmeans::contrast(model_mcgs_depth_species.i1, method='pairwise')

model_mcgs_depth_species.i2 <- emmeans::emmeans(model_mcgs_depth_species, ~ species | label | hemi)
emmeans::contrast(model_mcgs_depth_species.i2, method='pairwise')
```

## Plots: Depth (Figure 9a)
```{r}
# Figure 9a

# Manipulation for plot
mcgs_species_morphology.depth$hemi <- gsub("lh", "Left Hemisphere", mcgs_species_morphology.depth$hemi)
mcgs_species_morphology.depth$hemi <- gsub("rh", "Right Hemisphere", mcgs_species_morphology.depth$hemi)

mcgs_species_morphology.depth$label <- gsub("mcgs_db", "db", mcgs_species_morphology.depth$label)
mcgs_species_morphology.depth$label <- gsub("mcgs_mb", "mb", mcgs_species_morphology.depth$label)
mcgs_species_morphology.depth$label <- gsub("mcgs_sb", "sb", mcgs_species_morphology.depth$label)

mcgs_species_morphology.depth$label <- factor(mcgs_species_morphology.depth$label, 
                                              levels = c("mb", "db", "sb")
                                              )

# Make plot
mcgs_species_morphology.depth.plot <- mcgs_species_morphology.depth %>% 
  ggplot(aes(x = label, y = normalized_depth*100, alpha = species, fill = label)) + 
  
  geom_split_violin(width = 1) + 
  
  geom_boxplot(
               position = position_dodge(width =.14),
               outlier.size = 0, outlier.alpha = 0, 
               width = .13) +
  
  scale_y_continuous(name = "sulcal depth (% max)") +
  
  facet_wrap(~ hemi, nrow = 2) + 
  
  scale_fill_manual(breaks = c("mb", "db", "sb"),
                    values = c("cyan", "#33FF00", "#8856a7"),
                    ) +
  
  scale_alpha_manual(breaks = c("chimpanzee", "human"),
                     values = c(0.3, 0.65)) +
  
  guides(fill = "none", alpha = "none") +
  
  scale_y_continuous(n.breaks  = 5, limits = c(0, 100)) +

  plot_theme2

# View plot
mcgs_species_morphology.depth.plot

# Save plot
# ggplot2::ggsave(filename = "Figure_9a_mcgs_species_morphology_depth.png",
#                  plot = mcgs_species_morphology.depth.plot,
#                  device = "png",
#                  width = 8,
#                  height = 5,
#                  units = "in",
#                  dpi = "retina")

```

## Statistics: Surface area
```{r}
# Analysis 6.2: Surface area

# Select necessary columns and sulci with incidence > 50%
mcgs_species_morphology.SA <- mcgs_species_morphology %>% 
  select(c("sub", "species", "hemi", "label", "total_surface_area_.mm.2."))

# Normalize surface area
mcgs_species_morphology.SA <- merge(mcgs_species_morphology.SA, lobe_morphology, by = c("sub", "hemi", "species"))
mcgs_species_morphology.SA <- mcgs_species_morphology.SA %>% mutate(normalized_SA = total_surface_area_.mm.2./cortex_sa)

# Remove NA values
mcgs_species_morphology.SA <- mcgs_species_morphology.SA %>% drop_na()

# Run analysis
model_mcgs_SA_species <- lme(normalized_SA ~ hemi*label*species, random = (~1|sub/hemi/label), data = mcgs_species_morphology.SA)

# View summary
model_mcgs_SA_species.aov <- anova(model_mcgs_SA_species)
model_mcgs_SA_species.aov

# Effect size
effectsize::eta_squared(model_mcgs_SA_species.aov)

# Digest main effects and interactions
model_mcgs_SA_species.m1 <- emmeans::emmeans(model_mcgs_SA_species, ~ label)
emmeans::contrast(model_mcgs_SA_species.m1, method='pairwise')

model_mcgs_SA_species.m2 <- emmeans::emmeans(model_mcgs_SA_species, ~ species)
emmeans::contrast(model_mcgs_SA_species.m2, method='pairwise')
```

## Plots: Surface area (Figure 9b)
```{r}
# Figure 9b

# Manipulation for plot
mcgs_species_morphology.SA$hemi <- gsub("lh", "Left Hemisphere", mcgs_species_morphology.SA$hemi)
mcgs_species_morphology.SA$hemi <- gsub("rh", "Right Hemisphere", mcgs_species_morphology.SA$hemi)

mcgs_species_morphology.SA$label <- gsub("mcgs_db", "db", mcgs_species_morphology.SA$label)
mcgs_species_morphology.SA$label <- gsub("mcgs_mb", "mb", mcgs_species_morphology.SA$label)
mcgs_species_morphology.SA$label <- gsub("mcgs_sb", "sb", mcgs_species_morphology.SA$label)

mcgs_species_morphology.SA$label <- factor(mcgs_species_morphology.SA$label, 
                                              levels = c("mb", "db", "sb")
                                              )

# Make plot
mcgs_species_morphology.SA.plot <- mcgs_species_morphology.SA %>% 
  ggplot(aes(x = label, y = normalized_SA*100, alpha = species, fill = label)) + 
  
  geom_split_violin(width = 1) + 
  
  geom_boxplot(
               position = position_dodge(width =.14),
               outlier.size = 0, outlier.alpha = 0, 
               width = .13) +
  
  scale_y_continuous(name = "surface area (% cortex)") +
  
  facet_wrap(~ hemi, nrow = 2) + 
  
  scale_fill_manual(breaks = c("mb", "db", "sb"),
                    values = c("cyan", "#33FF00", "#8856a7"),
                    ) +
  
  scale_alpha_manual(breaks = c("chimpanzee", "human"),
                     values = c(0.3, 0.65)) +
  
  guides(fill = "none", alpha = "none") +
  
  plot_theme2

# View plot
mcgs_species_morphology.SA.plot

# Save plot
# ggplot2::ggsave(filename = "Figure_9b_mcgs_species_morphology_surface_area.png",
#                  plot = mcgs_species_morphology.SA.plot,
#                  device = "png",
#                  width = 8,
#                  height = 5,
#                  units = "in",
#                  dpi = "retina")
```


# Side analyses not included 

## pos-cas analysis
```{r}
# Load dataframe
pos_cas <- read.csv("pos_cas.csv")

## Human
pos_cas_human <- pos_cas %>% subset(species == "human")

### View data
pos_cas_type.figure <- pos_cas_human %>% 
  ggplot(aes(x = as.factor(intersection_inflated))) + 
  geom_bar(aes(fill = as.factor(isms_pres)), color = "black", position = "dodge") 
pos_cas_type.figure

### Analysis
isms_pres_pos <- glm(isms_pres ~ as.factor(intersection_inflated)*hemi, 
                   data=pos_cas_human,
           family=binomial())
Anova(isms_pres_pos, test="LR", type="III")

## Chimpanzee
pos_cas_chimp <- pos_cas %>% subset(species == "chimpanzee")

### View data 1
pos_cas_type.figure2 <- pos_cas_chimp %>% 
  ggplot(aes(x = as.factor(intersection_inflated))) + 
  geom_bar(aes(fill = as.factor(isms_pres)), color = "black", position = "dodge") 
pos_cas_type.figure2

### View data 2
pos_cas_type.figure3 <- pos_cas %>% 
  subset(species == "chimpanzee") %>%
  ggplot(aes(x = as.factor(intersection_pial))) + 
  geom_bar(aes(fill = as.factor(isms_pres)), color = "black", position = "dodge") 
  #scale_y_continuous(name = "% of hemispheres", seq(0,100,20), limits = c(0,100)) 
pos_cas_type.figure3

### Analysis
isms_pres_pos.ch.pial <- glm(isms_pres ~ as.factor(intersection_pial)*hemi, 
                   data=pos_cas_chimp,
           family=binomial())
Anova(isms_pres_pos.ch.pial, test="LR", type="III")
```

